---
title: "Assignment 4 - Max Haberl"
date: "28/12/2020"
output:
  html_document:
    toc: True
author: Max Haberl (5407084) in collaboration with Lukas Guenner (5393972) (different codes)
runtime: shiny

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Assignment 4 - Multilevel linear models and dashboards with Shiny 
I hereby assure to have complied with the *Code of Conduct*


**Please  Note:** Copy the path of your directory including the datasets to the clipboard. The *wd()* function will paste it coherently to the working directory

```{r Preparation, echo = FALSE, error = TRUE}

# Example of the path I used and copied into the clipboard C:\Users\maxem\Documents\Uni\Master\WiSe 2021\DSPM 20\Assignments\Ass4
rm(list= ls()) # clearing workspace
wd <- function() {  # helps converting a copied file path into r-eligible strings, otherwise directly assign to path
  x <- readClipboard()
  x <- gsub("\\\\", "/", x)
  return(x)
  }
path <- wd() # sets path according to wd() fct
# path <- "YOUR_WORKING_PATH"
setwd(path)
getwd()
```

## Exercise 1 - Andrew Gelman’s radon example

Loading the data, merging and preprocessing as required.
```{r EX1}
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("dplyr")) install.packages("dplyr")

library(tidyverse)
library(dplyr)

household_level <- read.csv('household.csv')
county_level <- read.csv('county.csv')

full_set <- left_join(household_level, county_level, by = 'county')
clean_set <- full_set[,-5]
clean_set <- clean_set %>% rename(county.name = county.name.x)

```



## Exercise 2 - Varying-intercept model with an individual-level predictor

A quick glance at the frequencies of observations for each county shows that LAC QUI PARLE contains two household observations only. One observation being utterly implausible, it is not surprising that a FE-estimators of counties with very few observations only should not be trusted. Extreme values are more likely to occur for small sample cases. As for Hennepin yielding the largest absolute deviation the exact opposite holds true. With 105 housholds observed in this county, a lot more variation is present within this county and thus allows for more precision. This variability can be used and modelled in the multilevel setting.

```{r EX2}
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("lme4")) install.packages("lme4")

library(ggplot2)
library(lme4)

model1 <- lmer(log.radon ~ basement + (1 | county), data = clean_set) # estimating the first model
mixed1 <- coef(model1)$county[,1] # extracting the RE and FE coefficients

model2 <- lm(log.radon ~ basement + as.factor(county) - 1 , data = clean_set) # estimating model and remove intercept
no_pool2 <- coef(model2)[c(2:86)]

deviation1 <- abs(mixed1 - no_pool2) # absolute deviation betweeen the two
result1 <- cbind.data.frame(deviation1, county_level$county.name) # merge into one 
minmax1 <- result1 %>% subset(deviation1 == max(result1$deviation1) | deviation1 == min(result1$deviation1)) # look for max and min
test <- household_level %>% group_by(county.name) %>% summarise(n=n())
print(minmax1)

```

## Exercise 3 - Varying-intercept model with an individual-level and a group-level predictor

```{r EX3}
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("lme4")) install.packages("lme4")
if (!require("arm")) install.packages("arm")

library(ggplot2)
library(lme4)
library(arm)


model3 <- lmer(log.radon ~ basement + log.uranium + (1 | county) , data = clean_set)
mixed3 <- coef(model3)$county[,1] + coef(model3)$county[,3]*county_level$log.uranium + coef(model3)$county[,2] # error in study? why include the low-level predictor coefficient as well?
se_intercepts <- se.ranef(model3)$county
plot_data <- cbind.data.frame(county_level$log.uranium, mixed3, se_intercepts)

int_fix <- fixef(model3)[1]+fixef(model3)[2]
slope_fix <- fixef(model3)[3]



plot1 <- ggplot(plot_data, aes(x = county_level$log.uranium, y = mixed3)) +
  geom_point() + 
  geom_errorbar(aes(ymin = mixed3 - se_intercepts, ymax = mixed3 + se_intercepts )) +
  geom_abline(intercept =  int_fix, slope = slope_fix) + labs(x = 'county log(uranium) level', y = 'estimated county intercept', title = 'Visualization of 2nd level regression')

plot1
```

## Exercise 3 - Varying intercepts and slopes model

```{r EX4}

model4 <- lmer(log.radon ~ 1 + basement + (1 + basement | county), data = clean_set)
int_model4 <- coef(model4)$county[1]
slope_model4 <- coef(model4)$county[2]
helper <- cbind.data.frame(int_model4, slope_model4, county_level$county.name)

int_model1 <- coef(model1)$county[1]
slope_model1 <- coef(model1)$county[2]
helper2 <- cbind.data.frame(int_model1, slope_model1, county_level$county.name)


if (!require("shiny")) install.packages("shiny")

library(shiny)


ui <- navbarPage(title = 'Assignment IV Dashboard - Multilevel Modelling',
                 tabPanel('Model 1', sidebarLayout(
    sidebarPanel(
      selectInput(inputId = "x1", label = "County", choices = county_level$county.name, selected = "ANOKA")
    ),
    mainPanel(
      verticalLayout(textOutput('title1'),
                     plotOutput(outputId = "scat1"),
                     htmlOutput('text1'))))),
                 tabPanel('Model 2', 
                          verticalLayout(textOutput('title2'),
                                         plotOutput('scat2'))),
                 tabPanel('Model 3',   sidebarLayout(
    sidebarPanel(
      selectInput(inputId = "x", label = "County", choices = county_level$county.name, selected = "ANOKA")
    ),
    mainPanel(
      verticalLayout(textOutput('title'),
                     plotOutput(outputId = "scat"),
                     htmlOutput('text'))
    )
  ))

)



server <- function(input, output, session) {
    observeEvent(input$x, {
      dataset <- filter(clean_set, county.name == input$x)
      temp <- filter(helper, county_level$county.name == input$x)
      # dataset$intcpt <- rep(temp[1], length(dataset))
      # dataset$slpe <- rep(temp[2], length(dataset)) 
      intcpt <- as.numeric(temp[1])
      slpe <- as.numeric(temp[2])
      
      output$title <- renderText('Varying intercepts and slopes model')
      output$scat <- renderPlot({
        ggplot(dataset, aes(x = basement, y = log.radon)) + geom_abline(intercept = intcpt, slope = slpe) + 
          geom_point(position = "jitter", color = 'blue') + labs(x = 'basement indicator', y = 'log(radon) concentration', title = 'Multilevel Regression Lines')
      })
      
      output$text <- renderText({
        paste('For county <b>', input$x, '</b> the regression intercept is: <b>', round(intcpt, 4), '</b>. The slope is : <b>', round(slpe, 4),'</b> .', sep = '')
      })
  
    })
  
    observeEvent(input$x1, {
      dataset2 <- filter(clean_set, county.name == input$x1)
      temp <- filter(helper2, county_level$county.name == input$x1)
      # dataset$intcpt <- rep(temp[1], length(dataset))
      # dataset$slpe <- rep(temp[2], length(dataset)) 
      intcpt <- as.numeric(temp[1])
      slpe <- as.numeric(temp[2])
      
      output$title1 <- renderText('Varying-intercept model with an individual-level predictor')
      output$scat1 <- renderPlot({
        ggplot(dataset2, aes(x = basement, y = log.radon)) + geom_abline(intercept = intcpt, slope = slpe) + 
          geom_point(position = "jitter", color = 'blue') + labs(x = 'basement indicator', y = 'log(radon) concentration', title = 'Multilevel Regression Lines')
      })
      
      output$text1 <- renderText({
        paste('For county <b>', input$x1, '</b> the regression intercept is: <b>', round(intcpt, 4), '</b>. The slope is : <b>', round(slpe, 4),'</b> .', sep = '')
      })
  
    })
    
    output$title2 <- renderText('Varying-intercept model with an individual-level and a group-level predictor')
    output$scat2 <- renderPlot({
      ggplot(plot_data, aes(x = county_level$log.uranium, y = mixed3)) +
  geom_point(color = 'blue', size = 2) + 
  geom_errorbar(aes(ymin = mixed3 - se_intercepts, ymax = mixed3 + se_intercepts ), color = 'lightblue') +
  geom_abline(intercept =  int_fix, slope = slope_fix) + labs(x = 'county log(uranium) level', y = 'estimated county intercept', title = 'Visualization of 2nd level regression', subtitle = 'including 1 SE deviations')
    })
    
  
}

shinyApp(ui = ui, server = server)
```
